### 2023-08-03 analysis/supplementary_materials/map backgrounds.Rmd

---
title: "prepare map backgrounds"
author: "Nikolas Gestrich"
date: '2022-05-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rnaturalearth)
library(rnaturalearthdata)
library(dplyr)
library(ggplot2)
library(sf)
```

This document is intended to make transparent the processes behind the creation of maps in the article associated with this repository. While this is not strictly necessary for scientific reproducibility, it might nevertheless be interesting to some readers.

# Country and continent inset data

## Continent data

For the outline of Africa, we import a shapefile

```{r}
africa_outline <- sf::st_read(here::here("analysis", "data",
                                         "raw_data", "Africa_outline.shp")) |>
  sf::st_crop(xmax = 54, xmin = -20, ymax = 40, ymin = -36)
```

## country outline

Here, we subset a free dataset of administrative boundaries downloaded from naturalearthdata.com.

```{r subset adm0}
mali_adm0 <- ne_countries(scale = "medium", returnclass = "sf") |>
  filter(admin == "Mali")
```

These can be used for inset maps inset following manner

```{r plot adm0}
ggplot() +
  geom_sf(data = mali_adm0) +
  geom_sf_text(data = mali_adm0, aes(label = admin))

ggplot(africa_outline) +
  geom_sf()
```

# Hydrology

```{r}
rivers <- ne_download(scale = 50, type = "rivers_lake_centerlines",
                      category = "physical", returnclass = "sf") |>
  sf::st_intersection(mali_adm0) |> filter(name == "Niger")

```

```{r plot hydro}
ggplot() +
  geom_sf(data = mali_adm0) +
  geom_sf_text(data = mali_adm0, aes(label = admin)) +
  geom_sf(data = rivers, fill = '#9ecae1') +
  geom_sf(data = osm_lakes.sf)
```

# Relief

## DEM

```{r}
library(raster)
#Mali outline
mali_sp <- sf::as_Spatial(mali_adm0)
#get SRTM elevation data
srtm1 <- getData("SRTM", lat = 14, lon = -3, download = TRUE)
srtm2 <- getData("SRTM", lat = 16, lon = -3, download = TRUE)
srtm3 <- getData("SRTM", lat = 14, lon = -6, download = TRUE)
srtm4 <- getData("SRTM", lat = 16, lon = -6, download = TRUE)
srtm5 <- getData("SRTM", lat = 22, lon = -6, download = TRUE)
srtm6 <- getData("SRTM", lat = 22, lon = -3, download = TRUE)
srtm7 <- getData("SRTM", lat = 18, lon = 2, download = TRUE)
srtm8 <- getData("SRTM", lat = 14, lon = -12, download = TRUE)
srtm9 <- getData("SRTM", lat = 16, lon = -12, download = TRUE)
srtm10 <- getData("SRTM", lat = 22, lon = 2, download = TRUE)

#Mosaic/merge srtm tiles
srtmmosaic <- mosaic(srtm1, srtm2, srtm3, srtm4, srtm5, srtm6, srtm7, srtm8, srtm9, srtm10, fun = mean)
#crop to Mali
dem1 <- crop(srtmmosaic, extent(mali_sp))
mali_srtm <- mask(dem1, mali_sp)
#downsample
mali_srtm_a <- terra::aggregate(mali_srtm, fact = 10)
#convert to df
dem.df <- rasterToPoints(mali_srtm_a)
dem.df <- data.frame(dem.df)
colnames(dem.df) <- c("lon", "lat", "alt")
```

## Hillshade

```{r}
# extract slope and aspect
slope.raster <- terrain(mali_srtm_a, opt = 'slope')
aspect.raster <- terrain(mali_srtm_a, opt = 'aspect')
# calculate hill shade
hill.raster <- hillShade(slope.raster, aspect.raster, 40, 270)
# convert to df
hill.m <- rasterToPoints(hill.raster)
hill.df <-  data.frame(hill.m)
colnames(hill.df) <- c("lon", "lat", "hill")
```

```{r}
ggplot() +
  geom_raster(data = hill.df, aes(lon, lat, fill = hill)) +
  scale_fill_gradientn(colours = grey.colors(100)) +
  geom_raster(data = dem.df, aes(lon, lat, fill = alt), alpha = 0.8) +
  scale_fill_gradientn(colours = terrain.colors(100)) +
  geom_sf_text(data = mali_adm0, aes(label = admin)) +
  geom_sf(data = rivers, colour = "blue") +
  theme_minimal()

```

# Save data
```{r save geodata}
save(africa_outline, mali_adm0, rivers, dem.df, hill.df,
     file = here::here("analysis", "data", "derived_data", "geodata.rda"))
```

From this data, we create several thematic map backgrounds for use in the article. These are stored in the `/R` folder, and are called to the paper file as source scripts.