### 2024-05-29 analysis/paper/paper.Rmd
changes accroding to paper revisions

---
title: "Pottery decor as networks on the Middle Niger"
author:
  - Nikolas Gestrich:
      email: gestrich@uni-frankfurt.de
      institute: [FI]
      correspondence: true
  - Juan-Marco Puerta-Schardt:
      email: Puerta-Schardt@em.uni-frankfurt.de
      institute: [GU]
      correspondence: false
institute:
  - FI: Frobenius Institute
  - GU: Goethe University
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      #pandoc_args:
      #- --lua-filter=../templates/scholarly-metadata.lua
      #- --lua-filter=../templates/author-info-blocks.lua
      #- --lua-filter=../templates/pagebreak.lua
abstract: |
  This paper reports on an attempt to analyse decorative techniques on archaeological pottery from the Middle Niger as knowledge exchange networks. The 12th-century CE state of this network is modelled and analysed, and then compared to its 9th-century counterpart. The analysis shows how knowledge about pottery decoration was propagated within networks of different sizes and intensities, from local, well-connected networks to more loosely tied long-distance connections. The article shows how the perspective of network analysis differs from, and can fruitfully complement, previous treatments of ceramic décor distribution in the region. It adds a novel perspective to what the distribution of archaeological ceramics in this region reflects, and contributes to the generation of hypotheses that can be further tested by fieldwork. 
keywords: |
  Mali; network analysis; pottery; archaeological ceramics
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r knitr-setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

```{r libraries, echo = FALSE}
# libraries

# Load the dplyr package for data manipulation
library(dplyr)

# Load the knitr package for generating reports using RMarkdown
library(knitr)

# Load the patchwork package for combining multiple plots into a single plot
library(patchwork)

# Load the ggplot2 package for creating graphics and plots
library(ggplot2)

# Load the tidygraph package for working with tidygraph objects
library(tidygraph)

# Load the ggraph package for visualizing graph data
library(ggraph)

# Load the tibble package for creating and manipulating tabular data
library(tibble)

# Load the igraph package for working with network data
library(igraph)

# Load the tidyr package for data tidying and reshaping
library(tidyr)
library(tidyverse)

# Load the sf package for reading, writing, and manipulating spatial data
library(sf)

# Load the multinet package for analysis of multilayer networks
library(multinet)

# Load the kableExtra package for nicer table design
library(kableExtra)

# Load the readr package to read excel-data
library(readr)
```

<!-- The actual document text starts here: -->

The full text of the article is available in the published version in "Azania: archaeological research in Africa". This document contains the code for all analyses and figures in the text,as well as some supplementary analzses and figures. It is structured by section to ease overview.

For details on the data and its preparation, see the document in the folder "analysis/supplementary_materials".

# Introduction

## Middle Niger Ceramics:

### Lack of temporal markers:

### Ceramics and social boundaries:

### Ceramics as trade goods:

## Ceramic decors as networks

# Methods

```{r import 11c decor data, echo = FALSE}
library(readr)
# import decor data

decors <- read_csv2(here::here("analysis/data/raw_data/decors12.csv")) |>
  select(-1) |> 
  rename(sites = Site) |> 
  select(-slip, -unidentified_roulette, -"twisted_cord_roulette?")

decorsearly <-  read_csv2(here::here("analysis/data/raw_data/decors9.csv")) |>
  select(-1) |> 
  rename(sites = Site) |> 
  select(-slip, -unidentified_roulette)

# import information on sites
sites <- read_csv2(here::here("analysis/data/raw_data/sites_all3.csv"))

# import decor count per site
n_decors <- read_csv2(here::here("analysis/data/raw_data/n_decor.csv")) |> 
  rename(sites = site) |> 
  select(-1)
```

## Data

```{r sitemap, fig.cap="location of sites in the analysis", echo = FALSE}
source(here::here("analysis/R/map plots.R")) # Load the required R script

#saving site location as sf
sitepoints <- st_as_sf(sites, coords = c("lon", "lat"),
                       crs = 4326, agr = "constant")

#creating a map of mali with the sites
sitemap <- ggplot() +
  map_bw_grid +
  geom_sf(data = sitepoints)

#ploting map with inset
sitemap + inset_element(inset, 0.72, 0.72, 1, 1)
```

## Network construction

## Data Quality

```{r iNext overview, echo = FALSE}
library(iNEXT)

#getting absolute numbers of decorations for the 12th century sites
decors_abs <- decors |> 
  left_join(n_decors) |> 
    mutate(across(where(is.numeric), ~ .*n/100)) |> 
  mutate(across(where(is.numeric), ~ round(.))) |> 
    column_to_rownames("sites") |> 
  select(-n) |> 
t()

# getting a table of the results of the analysis
i <- iNEXT(decors_abs, q=0, datatype="abundance")

i

# Fractured and extrapolated diversity of Doupwil
  i <- decors_abs[,c(4)]|> 
 iNEXT(q=c(0), datatype="abundance")
 ggiNEXT(i, type=1, facet.var="Assemblage")+
  labs(title = "Doupwil")

# Fractured and extrapolated diversity of Jenne-jeno
 i <- decors_abs[,c(8)]|> 
 iNEXT(q=c(0), datatype="abundance")
  ggiNEXT(i, type=1, facet.var="Assemblage")+
  labs(title = "Jenne-jeno")

#getting absolute numbers of decorations for the 9th century sites
decors_abs <- decorsearly |> 
  left_join(n_decors) |> 
    mutate(across(where(is.numeric), ~ .*n/100)) |> 
  mutate(across(where(is.numeric), ~ round(.))) |> 
    column_to_rownames("sites") |> 
  select(-n) |> 
t()

# getting a table of the results of the analysis
iNEXT(decors_abs, q=0, datatype="abundance")
```

# Analysis

## Multiplex structure

```{r make decor network, echo = FALSE}
# import functions for network construction
source(here::here("analysis/R/edgelist functions.R"))

# import custom functions for network analysis
source(here::here("analysis/R/network functions.R"))

# import custom functions for data visualisation
source(here::here("analysis/R/datavis themes.R"))

# apply function to transform frequency table into edge list
edgelist <- make_edgelist(decors)

# construct network, remove extraneous sites
decor_graph <- tbl_graph(sites, edgelist, directed = FALSE) |>
  activate(nodes) |>
  filter(!node_is_isolated())
```

```{r total-network, fig.cap="Plot of 11th century AD network of decorative practices. The plot shows the very strongly connected structure of the network."}
# Creating a multi-layer graph
ggraph(decor_graph, layout = "stress") +
  geom_edge_fan(colour = "grey") +
  geom_node_point()+
  geom_node_text(aes(label = label))+
  theme_graph()
```

```{r total-network-sample, fig.cap="Partially overlapping networks of four decorative practices", echo = FALSE}
decor_graph |>
  activate(edges) |>
  # Filter "edges" based on 4 specific decorations
  filter(name %in% c("braided_cord_roulette",
                     "braided_strip_roulette",
                     "paint",
                     "straight_mat_impression")) |>

  #plot the graph
  ggraph(layout = "stress") +
  coloured_edges +
  theme_graph()
```

```{r import to multinet, echo = FALSE}
library(multinet)
# reduce graph to types with over 5 edges
good_edges <- decor_graph |>
  activate(edges) |>
  as_tibble() |>
  count(name) |>
  filter(n > 5)

#creating a reduced graph with only those edges
reduced_graph <- decor_graph |>
  activate(edges) |>
  filter(name %in% good_edges$name)

# prepare graph for multinet by renaming
prep_graph <- reduced_graph |>
  activate(edges) |>
  rename(layer = name) |>
  activate(nodes) |>
  rename(name = label)

#make a vector of different layer names
layers <- prep_graph |>
  activate(edges) |>
  as_tibble() |>
  distinct(layer) |>
  pull(layer)

# get a list of individual networks using
#the onelayer function (/R/network functions.R)
graphlist <- purrr::map(layers, onelayer)
names(graphlist) <- layers

# add layers into a multinet object
multi <- ml_empty()

for (i in seq_along(layers)) {
  add_igraph_layer_ml(multi, graphlist[[i]], paste(layers[[i]]))
}
```

```{r layer-comparison, fig.cap="Jaccard similarity of nodes present in the networks of individual decorative practices"}
# perform layer compariosn using jaccard similarity between nodes
lc <- layer_comparison_ml(multi, method = "jaccard.actors")

# make heatmap
heatmap_matrix(lc)
```

## Sparsification

### Simplified network backbone

```{r flatten}
decor_flat <- flatten(decor_graph, weighted = TRUE) |> 
activate(nodes)

```

```{r backbone-extraction, fig.cap="Extraction of binary backbones at decreasing significance levels: a) 0.05, b) 0.15, c) 0.18"}
# extract binary backbones at decreasing levels of significance
decor_05 <- extract_backbone(decor_flat, 0.05)
decor_15 <- extract_backbone(decor_flat, 0.14)
decor_20 <- extract_backbone(decor_flat, 0.20) #20 ideale anzahl?

# plot
p1 <- plot.bb(decor_05)
p2 <- plot.bb(decor_15)
p3 <- plot.bb(decor_20) + geom_node_text(aes(label = label), repel = TRUE)

(p1 | p2) / p3 +
  plot_annotation(tag_levels = "a") +
  plot_layout(heights = unit(c(1, 8), c("cm", "null")))

decor_20 <- decor_20 |> 
  #isolated nodes are removed
  activate(nodes) |>
  filter(!node_is_isolated())

```

```{r weighted flat decor}
flat_edges <- decor_flat |> 
  get.data.frame("edges")

decor_20 <- decor_20 |> 
  activate(edges) |> 
  select(-weight) |> 
  left_join(flat_edges)

ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
  geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

### Influential nodes

```{r backbone-eigenvalue, fig.cap="Binary backbone at alpha = 0.15 showing eigenvector centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_eigen())) +
   geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "eigenvector centrality")
```

### Brokers

```{r backbone-betweenness, fig.cap="Binary backbone at alpha = 0.15 showing betweenness centralities"}
ggraph(decor_20, layout = "nicely") +
  geom_node_point(aes(size = centrality_betweenness())) +
   geom_edge_fan(color = "grey", alpha = 0.8, aes(edge_width = weight)) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph() +
  labs(size = "betweenness centrality")
```

### Evaluating measurements

```{r validity-test, warning = FALSE, message = FALSE, fig.cap= "Testing the robustness of the bb-network by comparing its centralities to the original graph"}
#comparing
p1 <- test_centralities(decor_graph, "eigen", 100)
p2 <- test_centralities(decor_graph, "betweenness", 100)

(p1 | p2) + plot_annotation(tag_levels = "a") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Community detection

```{r community detection, hide = 'results', echo=FALSE}

#sparsifying 
decor_graph_c <- decor_20 |>
  activate(nodes) |>
  mutate(louvain = as.character(group_louvain()))

decor_graph <- decor_graph |>
  activate(nodes) |>
  mutate(louvain = as.data.frame(decor_graph_c, "vertices")$louvain)

# calculate modularity of louvain clustering
modularity <- decor_20 |>
  activate(nodes) |>
  mutate(modularity = graph_modularity(group = group_louvain())) |>
  pull(modularity) |>
  head(1)
```

```{r plot-clusters, fig.cap="Louvain communities within the decor network", echo = FALSE}
# set plot colours for communities
communitycolor <-
  setNames(
    c("cornflowerblue","tomato", "darkgreen", "yellow"),
    unique(V(decor_graph)$louvain)
  )

# creating a palette with the proper amount of colors
col_21 <- colorRampPalette(brewer.pal(8, "Set2"))(21)

# then we create a vector containing all decortypes
decortypes <- decor_graph %>%
  activate(edges) %>%
  as_data_frame(what = "edges") %>%
  select(name) %>%
  unique()

decortypes <- as.vector(decortypes[, 1])

# finally we make a named vector of the colors
decorcolor <- setNames(col_21, decortypes)

#plot:
# disable s2 geometry to plot on map
sf_use_s2(FALSE)

ggraph(decor_graph, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat,
                              fill = louvain, colour = louvain),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor) +
  scale_colour_manual(values = communitycolor) +
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()

```

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  filter(label != "Sanga D") |> 
  activate(edges) |> 
  #filter(weight > 3) |> 
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.5, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
 flatten(weighted = TRUE) |> 
 activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
flatten(weighted = TRUE) |> 
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.5, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()

decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(x = lon, y = lat) +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

```{r cluster2, fig.cap="Internal connections within community 2"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "tomato") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

```{r cluster4, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

```{r cluster-comparison, fig.cap="Relative edge weights, edge diversity and centralization by cluster"}
# adding community/crossing to the edges
decor_graph <- decor_graph |>
 # flatten(weighted = TRUE) |> 
  add_community_edges()

# mean edge weights
mean_edge_weight <- decor_graph |>
  get.data.frame("edges") |>
#  filter(community != "crossing") |>
  group_by(from,to, community) |> 
  summarise(weight = sum(weight),
  div = n()) #|> 
 # group_by(community) |>
 # summarise(median_weight = median(weight),
  #          mean_weight = mean(weight))

# Centralization of the cluster
centralization <- vector("double", length(4))
for (i in 1:4) {
  centralization[[i]] <- decor_graph |>
    # filter community
    activate(nodes) |>
    filter(louvain == i) |>
    # degree centralization
    centr_degree(mode = "all", normalized = TRUE) |>
    as_tibble() |>
    pull(centralization) |>
    head(1)
}

# Graph of mean edge weight of the different communities
g1 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = weight, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "mean edge weight",
       x = "community")

# Graph of edge diversity of the different communities
g2 <- ggplot(mean_edge_weight) +
  geom_boxplot(aes(x = community, y = div, fill = community)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "edge div",
       x = "community")

#Graph of centralization of the communities
g3 <- tibble(centralization) |>
  mutate(community = 1:4) |>
  pivot_longer(!community, names_to = "key", values_to = "centralisation") |>
  ggplot(aes(x = community, y = centralisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw()

# plot:
g1 + g2 +g3
```

```{r cluster_centralities, fig.cap="Distribution of weighted degree and eigenvector centralities by community."}

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph)

# Display boxplots of centralities for degree centrality
a <- ggplot(centr) +
  geom_boxplot(aes(x = louvain, y = deg, fill = louvain)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "degree centrality",
       x = "community")

# Display boxplots of centralities for eigenvector centrality
b <- ggplot(centr) +
  geom_boxplot(aes(x = louvain, y = eig, fill = louvain)) +
  scale_fill_grey(start = 0.5, end  = 1) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(y = "eigenvector centrality",
       x = "community")

# Combine the boxplots
a + b
```

```{r crossings, fig.cap="Network of edges which cross between clusters"}

#show crossings with a weight over 3
decor_graph |>
  filter(
    community == "crossing" &
    weight > 3 
  # & !name %in% c("channel", "twisted_cord_roulette")
  ) |>
  activate(nodes) |>
  filter(!node_is_isolated()) |>
  #plot
  ggraph(layout = "stress") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight),
                      color = louvain), show.legend = FALSE) +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  scale_color_manual(values = communitycolor) +
  theme_graph() +
  theme_community
```

## Diachronic analysis

```{r import-new-data}
library(readr)
decorsearly <- 
  read_csv2(here::here("analysis/data/raw_data/decors9.csv")) |>
  select(-1)

# apply function to transform frequency table into edge list
edgelist <- make_edgelist(decorsearly)

# construct network, remove extraneous sites
decor_graph_early <- tbl_graph(sites, edgelist, directed = FALSE) |> 
  activate(nodes) |>
  filter(!node_is_isolated())
```

```{r early-backbone, fig.cap="early backbone"}
# flatten with weights
early_flat <- flatten(decor_graph_early, weighted = TRUE)

library(backbone)
# extract backbone
early_flat_bb <- 
  backbone::disparity(as.igraph(early_flat), alpha = 0.21, narrative = TRUE) |>
  tidygraph::as_tbl_graph()

# igraph transformation loses the node attributes, so we put them back
early_flat_nodes <- early_flat |>
  activate(nodes) |>
  as_tibble()
early_flat_bb <- early_flat_bb |>
  activate(nodes) |>
  mutate(label = early_flat_nodes$label,
         lat = early_flat_nodes$lat, 
         lon = early_flat_nodes$lon
         )

# put cumulative edge weights back into the backbone
flat_edges <- early_flat |> 
  get.data.frame("edges")

early_flat_bb <- early_flat_bb |> 
  activate(edges) |> 
  select(-weight) |> 
  left_join(flat_edges)

#plot
ggraph(early_flat_bb, layout = "stress") +
  geom_node_point(aes(size = centrality_betweenness())) +
  geom_edge_fan(aes(width = weight), color = "grey", alpha = 0.8) +
  scale_edge_width(range = c(0.2, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph()
```

# Discussion

# Conclusion

# Acknowledgements

The work presented here was undertaken as part of the project "Borrowed words and shared objects" within the German Research Foundation's (DFG) Priority Programme "Entangled Africa". Nikolas Gestrich was responsible for the research design and implementation. He is the main author of the analysis code and the manuscript. Juan-Marco Puerta Schardt aided in collating the data and co-authored the analysis code, particularly the section on community detection. He also contributed to the draft versions of the manuscript. The authors wish to thank the Institut des Sciences Humaines du Mali, the Musée National du Mali, and the Rijksmuseum voor Volkenkunde in Leiden for access to their pottery collections, and Susan McIntosh and Douglas Post Park for providing primary data. The R compendium for this article was created using the rrtools package.

<!-- The following line inserts a page break  -->

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
# if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at?
if ("git2r" %in% installed.packages() &
    git2r::in_repository(path = ".")) 
  git2r::repository(here::here()) 
```