### 2024-06-03 analysis/supplementary_materials/data preparation.qmd
last update inext

---
title: "Data preparation"
author: "Nikolas Gestrich and Juan Marco Puerta Schardt"
format: html
editor: visual
---

```{r libraries}
library(dplyr)
library(readxl)
library(readr)
library(tibble)
library(tidyr)
library(stringr)
```

## Jenné-jeno

For the 12^th^ century network, Jenne -jeno LXN levels 15-24 were used, which date to the period AD 1030-1200. For the 9^th^ century network, LXN levels 29-40 were used which date to the late 9^th^ early 10^th^ centuries (S. K. McIntosh pers.comm.) The raw sherd data were kindly given to us by Susan Keech McIntosh.

```{r jenne-jeno data prep 1}
# read data
Jj <- read_xlsx(here::here("analysis/data/raw_data/Jenne-jeno 1981 rim sherd data.xlsx"), sheet = 2)

# subset 12th century contexts
Jj12 <- Jj |> 
  filter(PIT == 1,
         Level > 14,
         Level < 25)

# subset 9th century contexts
Jj9 <- Jj |> 
  filter(PIT == 1,
         Level > 28,
         Level < 41)
```

Decors are coded separately in three columns, Twine, Paint, and Plastic. The slip coded in the surface columns is left our due to uneven reporting at other sites. We now prepare percentage tables for each of these, standardise their naming, and then merge them together.

```{r jenne-jeno data prep 2}
# Twine decors
Jj12twine <- Jj12 |> 
  count(Twine) |> 
  mutate(percent = n/143 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Twine, values_from = percent) |> 
  rename_with( ~ paste0("tw", .x))

# standardise nomenclature
Jj12twine <- Jj12twine |> 
  mutate(braided_cord_roulette = tw1 + tw2 + tw3, # Twine patterns 1, 2, and 3 are all made by rouletting a braided cord
         folded_strip_roulette = tw4 + tw5, # Twines 4 and 5 are folded strip roulette patterns
         twisted_cord_roulette = tw6 + tw7) |>  # twines 6 and 7 are made using a twisted cord roulette
  select(ends_with("roulette"))

# the code 40 in twine is unclear from the literature, and thus left out

# Paint decors
Jj12paint <- Jj12 |> 
  count(Paint) |> 
  mutate(percent = n/143 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Paint, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

Jj12paint <- Jj12paint |>  
  select(-x0)

Jj12paint <- Jj12paint |>    
  mutate(paint = rowSums(Jj12paint)) |> # summing all paint decors regardless of colour
  select(-starts_with("x"))

# Plastic decors
Jj12plastic <- Jj12 |> 
  count(Plastic) |> 
  mutate(percent = n/143 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Plastic, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

Jj12plastic <- Jj12plastic |> 
  mutate(channel = x1+x2, #single and multiple channels are treated as one category
         nonparallel_grooves = x3,
         impressed_comb = x4+x46+x47,
         finger_arcs = x13,
         dragged_comb = x24+x29+x93,
         punctate = x27, 
         applied_plastic = x91) |> 
  select(-starts_with("x"))

 Jj12decor <- bind_cols(Jj12paint, Jj12plastic, Jj12twine)  
 rownames(Jj12decor) <- "Jenne-jeno12"
```

Same procedure for the 9th century contexts

```{r jenne-jeno data prep 3}
# Twine decors
Jj9twine <- Jj9 |> 
  count(Twine) |> 
  mutate(percent = n/360 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Twine, values_from = percent) |> 
  rename_with( ~ paste0("tw", .x))

# standardise nomenclature
Jj9twine <- Jj9twine |> 
  mutate(braided_cord_roulette = tw1 + tw2 + tw3 + tw10, # Twine patterns 1, 2, 3, and 10 are all made by rouletting a braided cord
         folded_strip_roulette = tw4 + tw5, # Twines 4 and 5 are folded strip roulette patterns
         twisted_cord_roulette = tw6 + tw7 + tw12, # twines 6 and 7 are made using a twisted cord roulette
         cord_wrapped_stick_roulette = tw14) |> 
  select(ends_with("roulette"))

# the code 40 in twine is unclear, and thus left out

# Paint decors
Jj9paint <- Jj9 |> 
  count(Paint) |> 
  mutate(percent = n/360 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Paint, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

Jj9paint <- Jj9paint |> 
  select(-x0)

Jj9paint <- Jj9paint |> 
  mutate(paint = rowSums(Jj9paint)) |> # summing all paint decors regardless of colour
  select(-starts_with("x"))

# Plastic decors
Jj9plastic <- Jj9 |> 
  count(Plastic) |> 
  mutate(percent = n/360 * 100) |> 
  select(!n) |> 
  pivot_wider(names_from = Plastic, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

Jj9plastic <- Jj9plastic |> 
  mutate(channel = x1+x2+x18, #single and multiple grooves are treated as one category
         nonparallel_grooves = x3,
         finger_arcs = x13, 
         applied_plastic = x91,
         applied_ridges = x11,
         incised = x9+x12+x39,
         fingernail_impression = x18) |> 
  select(-starts_with("x"))

 Jj9decor <- bind_cols(Jj9paint, Jj9plastic, Jj9twine)  
 rownames(Jj9decor) <- "Jenne-jeno9"

```

Creating a dataframe with the number of decorations in Jenne-Jeno

```{r}
n_jj <- data.frame(
  site = c("Jenne-jeno9","Jenne-jeno12"),
  n = c(360,143)
)
```

## Togu Missiri

The data for the 12th century network includes rim and body sherds from two site phases, C and D (contexts 12 - 23), which comprised of several midden and two floor level deposits. They are dated with two radiocarbon dates to span a period from 1039 to 1222 cal AD.

```{r Togu data prep 1}
togu <- read_csv2(here::here("analysis/data/raw_data/TOG2 Rim sherds.csv"))

 # subset 12th century contexts
togu12 <- togu |>
  mutate(Context = as.numeric(Context)) |> 
  filter(Context > 11,
         Context < 24)
```

Decors are coded in four columns, beginning with "Motif". These are counted and their percentages of the assemblage calculated. Their naming is then standardised.

```{r Togu data prep 2}
# calculate percentages
togu_decor <- c(togu12$Motif1, togu12$Motif2, togu12$Motif3, togu12$Motif4) |> 
  as_tibble() |> 
  count(value) |> 
  mutate(percent = (n/148)*100) |> 
  select(-n) |> 
  pivot_wider(names_from = value, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

# standardise naming
togu_decor <- togu_decor |> 
  mutate(
    channel = xCH+xMCH,
    braided_cord_roulette = xCR2,
    folded_strip_roulette = xCR4+xCR4G,
    twisted_cord_roulette = xCR6,
    paint = `xP-R` + `xP-R-L` + `xP-R-L-N` + `xP-R-W` + `xP-W-L`,
    applied_plastic = xPA,
    applied_ridges = xPA3,
    fish_spine_roulette = xVPR
  ) |> 
  select(-starts_with("x"))

rownames(togu_decor) <- "Togu"

#number of decors in togu missiri
n_togu <- data.frame(
  site = c("Togu"),
  n = c(148)
)
```

## **Marakaduguba**

Finds used here come from an unpublished sequence in Unit C, excavated by Mamadou Cissé, Bourahima Ouedraogo and Nafogo Coulibaly. Levels 38, 39, and 40 are used, which consist of construction materials as well as a midden deposit. The levels are dated with two radiocarbon dates (from 40 and 38) to between 770 and 900 cal CE.

```{r MDB data prep 1}
mdb <- read_csv(here::here("analysis/data/raw_data/MDB_rims.csv"))

 # subset 9th century contexts
mdb9 <- mdb |>
  mutate(Context = as.numeric(Context)) |> 
  filter(Context %in% c(38,39,40))
```

Decors are coded in five columns, beginning with "Motif". These are counted and their percentages of the assemblage calculated. Their naming is then standardised.

```{r MDB data prep 2}
# calculate percentages 
mdb_decor <- c(mdb9$Motif1, mdb9$Motif2, mdb9$Motif3, mdb9$Motif4, mdb9$Motif5) |>    
  as_tibble() |>    
  count(value) |>    
  mutate(percent = (n/59)*100) |>    
  select(-n) |>    
  pivot_wider(names_from = value, values_from = percent) |>
  rename_with( ~ paste0("x", .x))

# standardise naming 
mdb_decor <- mdb_decor |>    
  mutate(     
    channel = xCH+xMCH,     
    folded_strip_roulette = xCR4,     
    twisted_cord_roulette = xCR6,     
    paint = `xPR`,     
    incised= xSI2+xSI4,     
    fish_spine_roulette = xVPR   
    ) |>   
  select(-starts_with("x"))

rownames(mdb_decor) <- "Marakaduguba"

#number of decorations in Marakaduguba
n_mdb <- data.frame(
  site = c("Marakaduguba"),
  n = c(59)
)
```

## Tongo Maaré Diabal

Sherds used in the 12^th^ century network stem from the upper levels of Horizon 5, which is dated to 1000-1150. By using only the uppermost layers of this Phase, we attempt to eliminate most of the 11^th^ century material from our sample. For the 9^th^ century network, sherds from Horizon 3 upper (750-900) are used.

```{r TMD data prep 1}
tmd <- read.csv2(here::here("analysis/data/raw_data/TMD93-96 decor.csv"))

# subset 9th and 12th century contexts
tmd12 <- tmd |> 
  filter(Phase == "5 up")

tmd9 <- tmd |> 
  filter(Phase == "3 up")
```

As at Togu Missiri and Duguba, the decorative elements are coded in columns prefixed with "Motif". These are extracted, percentages calculated, and names standardised.

```{r TMD data prep 2}
# 12th c.
# calculate percentages 
tmd12_decor <- c(tmd12$Motif1, tmd12$Motif2, tmd12$Motif3, tmd12$Motif4)|>    
  as_tibble() |> 
  mutate(value = if_else(value=="", NA_character_ ,value))|>    
  count(value) |>    
  mutate(percent = (n/143)*100) |>    
  select(-n) |>    
  pivot_wider(names_from = value, values_from = percent) |>
  rename_with( ~ paste0("x", .x))

# standardise naming 
tmd12_decor <- tmd12_decor |>    
  mutate(  
    applied_ridges = xAP3,
    braided_cord_roulette = xCCRY+xTW1,
    channel = xCH+xMCH, 
    cord_wrapped_stick_roulette = xCSR1+xCSR7,
    dragged_comb = xDC1+xDC2,
    dragged_organics = xGRASS,
    straight_mat_impression = xICS1+xICS4+xICS7,
    folded_strip_roulette = xTW4,     
    twisted_cord_roulette = xTW6+xTW12,     
    paint = `xP`+ `xP B`+`xP B-G`+`xP B-W`+`xP R`+`xP R-W`+`xP W`+   `xP W+B`+ `xP W-G` + `xP W-W`,     
    incised= xSI1+xSI3   
    ) |>   
  select(-starts_with("x"))

rownames(tmd12_decor) <- "TMD12"

# 9th century

# calculate percentages 
tmd9_decor <- c(tmd9$Motif1, tmd9$Motif2, tmd9$Motif3, tmd9$Motif4)|>    
  as_tibble() |> 
  mutate(value = if_else(value=="", NA_character_ ,value))|>    
  count(value) |>    
  mutate(percent = (n/73)*100) |>    
  select(-n) |>    
  pivot_wider(names_from = value, values_from = percent) |>
  rename_with( ~ paste0("x", .x))

# standardise naming 
tmd9_decor <- tmd9_decor |>    
  mutate(  
    braided_cord_roulette = xCCRX+xCCRY,
    channel = xCH+xMCH, 
    cord_wrapped_stick_roulette = xCSR1+xCSR7+xCSR8+xTW14,
    dragged_comb = xDC2,
    dragged_organics = xGRASS,
    straight_mat_impression = xICS1+`xPD/O`,
    twisted_cord_impression = xTI1,
    folded_strip_roulette = xTW4,     
    twisted_cord_roulette = xTW6+xTW12,     
    paint = `xP B`+`xP B-G`+`xP B-W`+`xP R`+`xP R-W`+`xP W`+`xP W-G` + `xP W-W`  
    ) |>   
  select(-starts_with("x"))

rownames(tmd9_decor) <- "TMD9"
```

```{r}
n_tmd <- data.frame(
  site = c("TMD9","TMD12"),
  n = c(73,143)
)
```

## Dia Mara

For the 12th century network, we have, by necessity, used body sherd data for Unit D6 Levels 6-8 (Haskell et al. 1988 Table 15 and approximated data from Figure 36), which McIntosh and McIntosh have recently reinterpreted as mostly falling into the 11th century (McIntosh et al. 2023, 21) with new AMS dates from levels 5 and 8 giving a combined span of 1026-1149 cal CE (McIntosh et al. 2023, 19, Table 7)

For the 9th century network, we have used the Horizon III data reported in Arazi (2005) Table 6.34 and 6.35. This is an agglomerate of rim and body sherd decors from Units M, S and Q. This is somewhat problematic as the Horizon III deposits in Unit M are undated, whereas those in Units Q and S are dated by a single date each (Bedaux et al. 2005; Arazi 2005). The date range for both is from the mid-7th to the mid-10th centuries (Q: GX-28750, 1250 +- 60; S: GX-30007-AMS, 1210+-40). Despite the fact that the main probablity ranges for both dates are in the 9th century, considerable insecurity over the stratigraphic and chronological integrity of this sample must remain.

```{r Mara data prep}
# 12th century
mara_12 <- read_csv2(here::here("analysis/data/raw_data/Mara 12.csv"))

n_mara12 <- mara_12 |> 
  filter(level == "total") |> 
  pull(n_total)

mara_12 <- mara_12|> 
  filter(level == "percent") |> 
  select(-c(n_total, level))

rownames(mara_12) <- "Mara12"

#renaming comb_impression
mara_12 <- mara_12 |> 
  rename(impressed_comb = comb_impression)
# 9th century
mara_9 <- read_csv2(here::here("analysis/data/raw_data/Mara 9.csv"))

#renaming comb_impression
mara_9 <- mara_9 |> 
  rename(impressed_comb = comb_impression,
         incised = incision)

rownames(mara_9) <- "Mara9"

#n of decorations at mara (number of Mara taken from Bedaux et al. 2005, 231)
n_mara <- data.frame(
  site = c("Mara9","Mara12"),
  n = c(281,512)
)
```

## ***Dia Shoma***

For the 12th century network, we use body sherd data from Unit Sh7 levels 6-9 (Haskell et al. 1988, Table 11 and Figure 34) which are dated by an AMS date from level 7 spanning from 1040-1165 cal CE (McIntosh et al. 2023, 19, Table 7). The only other date from occupation layers to fall into the 12th century range is from Unit C (1025-1159 cal CE.), but the corresponding ceramic data could not be isolated from the data published by Bedaux et al. or Arazi. The ceramics from these particular levels in Sh7 is discussed by McIntosh and McIntosh (2023) as being unusual and unknown in the remaining sequence at Dia, and as having clear affinities with the eastern IND. In including this material in the network, we should be aware of the fact that it might not well represent "typical" western IND pottery in this period.

For the 9th century, we did not feel comfortable using Horizon III ceramics, as reported in Bedaux et al. (2005). While most Horizon III dates from Shoma stem from burials, a further indication of the problematic nature of deposits from this period, Units F and B have dated Horizon III occupation layers whose dates span the entire second half of the first millennium CE, making the bracket too wide for meaningful comparison.

```{r Shoma data prep}
# 12th century
shoma_12 <- read_csv2(here::here("analysis/data/raw_data/Shoma 12.csv"))

n_Shoma12 <- shoma_12 |> 
    filter(level == "total") |> 
  rename(n = n_total) |> 
mutate(site = "Shoma12")|> 
select(site, n)

shoma_12 <- shoma_12|> 
  filter(level == "percent") |> 
  select(-c(n_total, level))

rownames(shoma_12) <- "Shoma12"
```

## **Essouk-Tadmekka**

The handmade and unglazed pottery data from Unit Ek-A Periods 3 a and 3b were used in the 12th century network \[\@nixonEssoukTadmekkaEarlyIslamic2017 Appendix E:21\]. It is not entirely certain whether this data refers to rims sherds only, or to rim and body sherds. The overall numbers of studied sherds in table E.21 does not correspond to any of the counts given in Table 12.2, and is somewhat higher than the numbers given for the total number of decorated sherds indicated there. It is therefore unclear what the sample represents. We assume that this an effect of multiple decors on some sherds and have taken the total number of decorated rim and body sherds as a basis for percentage calculation.

Despite the fact that Nixon (2017, 97) includes the 12th and 13th century in the dating for these phases, the radiocarbon dates for the deposits would rather indicate the 11th and 12th centuries as the date range. While only the basal layers of the preceding Period 2 are radiocarbon dated, the beginning of Period 3a has a *terminus post quem* of the 11th century from a sherd of Chinese Qingbai pottery \[\@nixonEssoukTadmekkaEarlyIslamic2017, p. 99-100\]. It therefore seems likely that the bulk of the deposits in these phases stem from the 12th century AD, with some possible admixture from the late 11th and early 13th centuries.

While there are 9th century deposits at Essouk-Tadmekka, they are either too insecurely dated (as in Unit Ek-B), or as in Unit Ek-A Periods 1 and 2, the pottery data presentation does not allow sufficient differentiation of the 9th century deposits from earlier and later ones.

```{r Essouk data prep}
essouk <- read_csv2(here::here("analysis/data/raw_data/Essouk Phase 3 pottery.csv")) |> 
  select(n,`decor description`) |> 
  mutate(percent = (n/450)*100) |> 
  select(-n) |> 
  pivot_wider(names_from = `decor description`, 
              values_from = percent, 
              values_fn = sum)|> 
  rename_with( ~ paste0("x", .x)) |> 
  mutate(across(everything(), as.numeric))

# standardise names
essouk <- essouk |>
  mutate(
    channel = `xsingle channel`+`xmultiple channels`,
    paint = `xpaint white cross-hatch`+
      `xpaint white line`+
      `xpaint whitetriangle`+
      `xpaint white wash`+
      `xpaint black line`+
      `xpaint black wash`+
      `xpaint red line`+
      `xpaint-red wash`,
    finger_impression = `xfinger impressions`+
      `xfingernail impression`, 
    dragged_comb = `xdragged comb linear` + 
      `xdragged comb geometric`,
    cord_wrapped_impression = `ximpressed cord-wrapped` +
      `ximpressed cord-wrapped stick`,
    incised_patterns = `xincised dashed lines`+
      `xincised geometric`+
      `xpunctate circular` +
      `xsideways stylus impression`
  ) |> 
  rename(
    twisted_cord_roulette = `xtwisted cord roulette`, 
    cord_wrapped_roulette = `xcord-wrapped roulette`
  ) |> 
  select(-starts_with("x"))

rownames(essouk) <- "Essouk"

#n of decorations in Essouk
n_ess <- data.frame(
  site = c("Essouk"),
  n = c(450)
)
```

## Sanga

While the cave sites studied by Bedaux were only sporadically dated, some contexts are secure enough to be useful for our analysis. This is in particular Grotte C, where very little disturbance was noted by the excavators, and where three dates on human bones in a collective burial of 170 individuals in the excavated 4,5 m2 have their main probability peaks in the late 11^th^ to late 13^th^ centuries. The two sigma range, however, extends from 979 to 1394. Given the nature of the site and the estimated total number of up to 3500 individual skeletons, a long use must be assumed. Grotte C is associated by the excavators with Grotte D below it, which are assumed to have been used at the same time.

For the 12^th^ century network, we have used ceramics from Grotte D recorded at the Rijksmuseum voor Volkenkunde in Leiden in 2012. We include this data despite its time range and problematic dating because the non-settlement nature of the site provides a counterpoint to the other assemblages. Due caution in the interpretation is necessary.

```{r tellem data prep}
tellem <- read_csv2(here::here("analysis/data/raw_data/LEIDEN_TELLEM.csv")) |> 
  filter(str_detect(Site, "^D"))

# calculate percentages
tellem_decors <- tellem |> 
  select(starts_with("Decor")) |> 
  pivot_longer(cols = everything()) |> 
  count(value) |> 
  mutate(percent = (n/57)*100) |> 
  select(-n) |> 
  pivot_wider(names_from = value, values_from = percent) |> 
  rename_with( ~ paste0("x", .x))

# standardise decors
tellem_decors <- tellem_decors |> 
  mutate(
    channel = xCH + xMCH,
    knotted_cord_roulette = xCR1,
    braided_cord_roulette = xCR2,
    twisted_cord_roulette = xCR6 + `xB4,CR6`,
    dragged_organics = xHERB,
    applied_ridges = xPA4,
    dragged_comb = xPE1,
    cord_wrapped_roulette = xPFR1 + xPFR2,
    incised_patterns = xSI4,
    braided_strip_roulette = xBSR
  ) |> 
  select(-starts_with("x"))

rownames(tellem_decors) <- "Sanga D"

#n of decorations in Essouk
n_sanga <- data.frame(
  site = c("Sanga"),
  n = c(57)
)
```

## Toguéré Doupwil

Sherds from section C, period k are used in the 12^th^ century network. The date GrN-7944 (1046-1275 cal CE) dates the lower layer 3 in section C. This layer consists of laminated sands, charcoal and pottery directly overlying the first evidence of building at the site in section C layer 2. This is summarised as period k in section C, which, also contains material from the underlying level 1, a thin occupational layer directly on floodplain material. While it is unlikely that level 2 is much older than the C14 date indicates, the relationship to level 1 is unclear. It is therefore possible that this sample contains material that significantly predates the 12th century.

In section B, Period e also includes the second half of the 12th century, as dated by GrN-7943 (1165-1283 cal CE), but the main peaks of this date fall in the 13th century and the sherds have therefore been excluded from the sample.

```{r doupwil data prep}
doupwil <- read_csv(here::here("analysis/data/raw_data/Toguere_Doupwil.csv")) |> 
  rename(twisted_cord_roulette = 'twisted_cord_roulette?')

rownames(doupwil) <- "Doupwil"

n_doupwil <- data.frame(
  site = c("Doupwil"),
  n = c(33)
)
```

## Toguéré Galia

The12^th^ century network includes pottery from Period II. This period, comprising of deposit groups C and D, is dated with a single C14 date to 1026--1217 cal CE (GrN-8555). This date was originally rejected by the excavators as being somewhat too late, since the uncalibrated BP range did not overlap with the underlying date for period I (GrN-7945). The latter date's calibrated range of 1038--1383 cal CE, however, removes this objection. Nevertheless, it must be cautioned that the sequence at this site is not securely dated.

```{r galia data prep}
#changing names to fit dataset
galia <- read_csv(here::here("analysis/data/raw_data/Toguere Galia.csv")) |> 
  rename(dragged_comb = drageed_comb,
         incised_patterns = incised_pattern)

rownames(galia) <- "Galia"

n_galia <- data.frame(
  site = c("Galia"),
  n = c(99)
)
```

## Akumbu

Stratigraphically detailed pottery décor data is available for the body sherds from units AK1 (Togola 2008, pp. 55--57) and AK5 (McIntosh 2017). The body sherd decor from AKB4 is reported in Horizons (Cissé et al. 2017, pp. 41--43). Décor data for rim sherds are not published for any of the excavations.

For the 12th century network, we have used selected data from Akumbu Mound A, Unit AK1's Middle Assemblage layers. In this unit, a living surface dated to 1266-1425 cal. CE (Beta-40441, 640 +- 70) was constructed above a refuse pit cut into mixed deposits, which the excavator believes to have accumulated slowly. The upper portion of the pit was dated to 1175-1329 cal. CE at 82.6% (Beta-40442, 750+-70). A further, undated, construction episode lay between these deposits and a curvilinear structure with a date (Beta-40443, 980+-70) of 954-1219 cal. CE at 92.7%. While the calibrated date range is rather wide, it was decided to use the pit levels (8,10,13,15) in our analysis. Unfortunately, the pottery data for these layers also includes layer 12, a wall collapse layer with uncertain stratigraphic relationship to the pit, rendering the dating of the sample more insecure. A sample more securely dated to the 12th century might have come from the upper levels of AK4, but the pottery was not recorded.

On Akumbu Mound B, the upper levels in the adjacent Units AK3 and AKB4 date to a range including the 9th century. However, the nearby AK5 has a thin 13th midden deposit overlying 3rd-4th century CE deposits. While the AK3 pottery is not described in detail, AKB4 Horizon 1 has two AMS dates, both from Level 2, which date to 771-976 cal. CE at 98.2% and 771-894 cal. CE at 88.3%. However, the nature of the levels is not made clear in the publication, most of the levels of Horizon 1 lie below this dated deposit, and it is unclear whether Horizon 1 includes the 5^th^-7^th^ century Level 8 or not (cf. Cissé et al. 2017, pp. 36 & 41). Trusting that this is not the case, as is also McIntosh's (McIntosh 2017, p. 197) interpretation, we have used the Horizon 1 body sherd data for the 9th century network.

There are problems with the pottery data, because paint and incisions are not listed separately even though they are mentioned at being present at the site of AKB4.

```{r}
#downloading the data from Cisse& Takezawa for the 9th century network
Akumbu9_decor <- read_csv2(here::here("analysis/data/raw_data/Akumbu early.csv")) |> rename("braided_cord_roulette" = "Cordelette roulee tressee en natte TW1",
          "folded_strip_roulette" = "Fibre plate pliee TW4",
"twisted_cord_roulette" = "Cordelette Torsadee TW6",
"dragged_comb" = "Peigne filete roule", #translation unsure
"slip" = "Engobe",
"other" = "Autre")

#getting the amount of sherds in Akumbu
n_Aku9 <- Akumbu9_decor |> 
  pull(n)

#removing undefinded sherds and n from table 
Aku9_decor <- Akumbu9_decor |> 
  select(-n, -other)|> 
  mutate(Site = "Akumbu9") |> 
  column_to_rownames("Site")

names(Aku9_decor)

```

```{r}
Akumbu12 <- read_csv2(here::here("analysis/data/raw_data/Akumbu_12.csv"))

#absolute numbers into proportion
Akumbu12_decor <- Akumbu12 |> 
  rename(Site = 1) |> 
column_to_rownames("Site")

  n_Aku12 <- Akumbu12_decor |> 
  select_if(is.numeric) |> 
  rowSums()

Aku12_decor <- Akumbu12_decor |> 
  add_column(n_Aku12) |> 
  mutate_if(is.numeric, function(x) x/n_Aku12*100) |> 
  select(-n_Aku12)

```

```{r}
n_akumbu <- data.frame(
  site = c("Akumbu9", "Akumbu12"),
  n = c(n_Aku9, n_Aku12)
)
```

## Ambere-Dougon

The site was excavated by Guindo (Guindo 2011, 243f). At the base of the uppermost layer, a piece of charcoal was found, dating to 1029-1209 cal. CE (Ly-3712, 915 ± 30 BP). We used the pottery from this Layer for our 12^th^-century network. The other date from the excavation comes from layer 5, the oldest of the excavation, dated to 430-599 cal. CE (Ly-3968, 1545 ± 30 BP). Unfortunately, the other layers cannot be precisely chronologically assigned, which makes it possible that one or more of these layers could correspond to our 9^th^-century network. However, for a reliable assignment, the corresponding radiocarbon d

During the recording of ceramic decorations, no distinction was made between red painting and slip. However, the author's descriptions lean more towards slip.

```{r}
AD <- read_csv2(here::here("analysis/data/raw_data/Ambere-Dougon.csv")) |> 
  rename(straight_mat_impression = "Impression de Vannerie",
         simple_braided_strip_roulette = "Impression de fibres plates tresses simple",
         alternate_braided_strip_roulette = "Impressions de Fibre plates tresses alterne",
        carved_wooden_roulette = "Cylindre grave",
        slip = "Engobe",
        twisted_cord_roulette = "Cordelette torsadee") |> 
  select(-Context) |> 
  mutate(Site = "Ambere-Dougon")  |> 
    column_to_rownames("Site") |> 
  mutate_all(as.numeric) |> 
  mutate(braided_strip_roulette = simple_braided_strip_roulette + alternate_braided_strip_roulette) |> 
  select(-slip, -simple_braided_strip_roulette, -alternate_braided_strip_roulette)

n_AD <- AD  |> 
  rowSums()

AD12_decor <- AD |> 
 mutate_if(is.numeric, function(x) x/33*100)

n_AD <- data.frame(
  site = c("Ambere-Dougon"),
  n = c(n_AD)
)

names(AD)
```

## Damassogou

The site was excavated by Néma Guindo (2011, 171). For the 12^th^-century network, pottery from Layer 2 was utilized. At the bottom of Layer 2 there is a structure that is only described as friable. The layer is dated to 925 ± 30 BP, corresponding to 1024-1206 AD44 (Ly 3189).

During the recording of ceramic decorations, no distinction was made between red painting and slip. However, the author's descriptions lean more towards slip.

```{r}
Dmssg <- read_csv2(here::here("analysis/data/raw_data/Damassogou.csv")) |> 
  rename(straight_mat_impression = "Impression de Vannerie",
         simple_braided_strip_roulette = "Impression de fibres plates tresses simple",
         alternate_braided_strip_roulette = "Impressions de Fibre plates tresses alterne",
        carved_wooden_roulette = "Cylindre grave",
        undecorated = "non decores",
        braided_cord_roulette = "Cordolette tressee simple",
        finger_impression = "Empreinte digitale",
        incised = "Incision")|> 
  mutate(Site = "Damassogou") |> 
    column_to_rownames("Site") |> 
  mutate_all(as.numeric) |> 
  mutate(braided_strip_roulette = simple_braided_strip_roulette + alternate_braided_strip_roulette) |> 
  select(-undecorated, -simple_braided_strip_roulette, -alternate_braided_strip_roulette)

n_Dmssg <- Dmssg  |> 
  rowSums()

Dmssg12_decor <- Dmssg |> 
 mutate_all(function(x) x/179*100)

names(Dmssg)

#n of pottery decors in Damassagou
n_Dmssg <- data.frame(
  site = c("Damassagou"),
  n = c(n_Dmssg)
)
```

## Gao

There have been several excavations at different sites at Gao; however, many of the older ones did not document ceramics in any detail (Cissé et al. 2013). Of the more recent excavations by Insoll (in 1993 & 1996) and the Malian Direction Nationale de Patrimoine Culturel (2001, 2002 & 2006), only the latter have documented ceramics in a way that can be used in our analysis. While we tried to re-record Insoll’s pottery which is stored at the Institut de Sciences Humaines in Bamako in 2020, selection errors meant that we did not have a suitably datable sample for this analysis. Our use of the ceramics reported by Cisse are complicated by the fact that there is an inconsistency between dates produced from different laboratories at the site, with those of the University of Waikato being consistently around 100 years older than the two related dates analysed by Beta Analytic (Cisse 2011:139). A further complication is added by Cisse's decision to present the ceramic data in three aggregated assemblages (Final, Upper, Lower) for each unit, making chronological control of the samples even more difficult. Furthermore, only pottery from ACGS can be used for our model as the roulette decorations have been sufficiently documented in this unit.

In Unit ACGS, the final levels (1-10) were made up of extensive trash deposits. Their upper component is dated to 995-1217 cal. CE (Beta 261360, 950 ± 60 BP). We include the "Final" assemblage in our twelfth century network as reflective of occupation at Gao, but there are points of caution: No architectural remains are known from this period at Gao Saney, the single date has a wide calibrated range and a date in neighbouring GS1, 40 cm lower, has a University of Waikato date with a calibrated range at the end of the first millennium CE. As only the top and the bottom of the ACGS sequence were dated, we cannot estimate how old the lower layers included in the sample are.

All of the units excavated by Cisse were probably inhabited during the ninth century, but to isolate the contexts is rather difficult. A simple sequence model in Oxcal 4.4 refines the date from GS3 Level 5 to 803-979 cal CE, and that from Level 12 to 772 - 939 cal CE, suggesting that the bulk of the deposits dates to the ninth and the first half of the 10th centuries. In the presentation of the ceramic data, however, the "Upper" section comprises of Levels 2-12 and its date range probably extends into the 11th century. In GS1, the "Upper" assemblage is narrower, uniting Levels 5-9. These are dated by a single date in Level 7b, to a modelled age of 696-884 CE (648-876 cal CE), suggesting an admixture of 8th and possibly even 7th-century materials. In order to limit the chronological range of the sample, only the Upper assemblage from GS1 has been used.

Although GS1 itself cannot be utilized, its proximity to ACGS enables a comparative analysis of the stratigraphy between the two units. Consequently, despite the absence of direct dating for the "Upper" Horizon of ACGS, we consider it the most suitable horizon for our ninth century network.

Both datasets contain cord roulette decorations that were not specified; 15.5% of the decorations documented for lower layers and 17.5% of the decorations documented for the final layers.

```{r}
Gao <- read_csv2(here::here("analysis/data/raw_data/Gao.csv")) |> rename("dragged_comb" = "Comb_tot",
    "channel" = "Channel_comb",
          "folded_strip_roulette" = "Folded strip roulette",
"twisted_cord_roulette" = "Twisted twine",
"paint" = "Paint_total", #translation unsure
"unidentified_roulette" = "Unident. Twine",
"cord_wrapped_roulette" = "Sisal", # Cisse (2011, 161) identifies them as cord_wrapped_roulette after Nixon (2008, 183)
"straight_mat_impression" = "Birdfood" # Cisse (2011, 161) identified it as mat, but not completly sure, maybe other mat than straight
) |> 
  filter(Phase != "lower") |> #removing the lowest phase as it is probably to old
  mutate(Site = "Gao") |> 
mutate_if(is.numeric, round, digits=0) #rounding the numbers (because they were extracted from graphics there were minor deviations)

Gao <- as.data.frame(Gao)

n_gao <- Gao |> 
  select_if(is.numeric) |> 
    rowSums()

Gao <- Gao |> 
  add_column(n_gao) |> 
  mutate_if(is.numeric, function(x) x/n_gao*100)

Gao9_decor <- Gao |> 
  filter(Phase == "upper") |> 
  select(-"Site-Code", -"Phase") |> 
  mutate(Site = "Gao9") |> 
  column_to_rownames("Site") |> 
  select(-"n_gao")

Gao12_decor <- Gao |> 
  filter(Phase == "final") |> 
  select(-"Site-Code", -"Phase", -"n_gao") |> 
  mutate(Site = "Gao12") |> 
  column_to_rownames("Site")

n_gao <- data.frame(
  site = c("Gao12", "Gao9"),
  n = c(n_gao)
)

Gao
```

## Kokolo NE

The Site was excavated by Keita (2011, 74). At the structure in Horizon 2 of the trench section A of Kokolo a radiocarbon date was taken (790 ± 35 BP 1190-1278 cal. CE). The author uses this date to contextualize horizons in other trenches to the same settlement Phase in the 12^th^ and 13^th^ century. Nevertheless, we will only use pottery from the dated horizon itself for our 12^th^ century network model.

```{r}
Kokolo <- read_csv2(here::here("analysis/data/raw_data/Kokolo NE.csv")) |> 
  rename(folded_strip_roulette = "fibres plates et pliees",
         twisted_cord_roulette = "cordelette torsadee",
         braided_cord_roulette = "tresse simple", #only as "l’impression à la tresse simple" in Keita 2011, pp. 33
        cord_wrapped_roulette = "roulette de cordelette enroulee")

n_Kokolo <- Kokolo |> 
  pull(n)

Kkl12 <- Kokolo |> 
  mutate(Site = "Kokolo")  |> 
  column_to_rownames("Site") |> 
  select(-n, -Context)

names(Kokolo)

n_kkl <- data.frame(
  site = c("Kokolo"),
  n = c(n_Kokolo)
)
```

## Nin-Bére

Nin-Bére was excavated by Guindo (2011, 227).From the Nin-Bèrè archaeological. site, there are two radiocarbon datings: one from Layer 2 dated to 730 ± 40 BP 1225-1376 cal. CE (Ly-13211) and another from Layer 5, which represents the latest excavation, dated to 1095 ± 35 BP 889-1016 cal. CE (Ly-13212). Since there are no structures providing a more precise chronological. placement for the other layers, only pottery from Layer 5 can be considered for inclusion in the 9^th^ century network.

During the recording of ceramic decorations, no distinction was made between red painting and slip. However, the author's descriptions lean more towards slip.

```{r}
Nin_Bere <- read_csv2(here::here("analysis/data/raw_data/Nin-Bére.csv")) |> 
  rename(folded_strip_roulette = "Impression de fibres plates plies",
         braided_cord_roulette = "Cordolette tressee simple",
         undefined = "Non determines") |> 
  mutate(Site = "Nin_Bere") |> 
  select(-Context) |> 
  column_to_rownames("Site")

n_NB <- Nin_Bere |> 
  rowSums()

Nin_Bere9 <- Nin_Bere |> 
    mutate_if(is.numeric, function(x) x/n_NB*100) |> 
  select(-undefined)

names(Nin_Bere)

n_NB <- data.frame(
  site = c("Nin-Bere"),
  n = c(n_NB)
)
```

## Promontoire site of Ounjougou

The excavation at the Promontoire site of the Ounjougou site complex was conducted by Downing as part of a diploma thesis. Ounjougou comprises seven levels, sometimes divisible into different horizons, with their summits appearing to correspond to occupation levels (Mayor et al., 2005). Level 4 is composed of two horizons. T4.2 was dated to 1315 ± 45 BP 660--720 cal. CE (Ly-1284 OXA), making it unsuitable for our 9^th^ century network. The other horizon of this level, T4.1, dates to 1005 ± 30 BP 990--1040 cal. CE (Ly-9340), thus being not beeintpart of our 12^th^ century network. We only used level T3, dating to 885 ± 50 BP 1040--1220 cal. CE (Ly-10150), for this network.

For our network models, we utilize pottery decoration data published by Mayor (2010). Her reassessment of the sherds aids in making them comparable to those from other sites within our model. But some decorations that are mentioned in other publications, like incisions will therefore not be taken into account.

```{r}
Ounjg <- read_csv2(here::here("analysis/data/raw_data/Promontoire Ounjougou.csv"))

#use T3 for the 12th century network and removing columns without values for this level
Ounjg12 <- Ounjg |> 
  filter(Level == "T3") |> 
  mutate(Site = "Ounjougou12") |> 
  column_to_rownames("Site") |> 
  select_if(is.numeric) |> 
    select_if(~all(. > 0)) |> 
  select(-"Non decore")

#renaming decorations
Ounjg12 <- Ounjg12 |> 
  rename(braided_strip_roulette = "Fibre plate tressee simple ou alterne",
         twisted_cord_roulette = "Cordelette torsadee",
         incised = "Incicions au couteau seules",
         channel = "Cordons" #this is described as "presénce d'un petit cordon sous un bord 
         )

#amount of sherds
n_ounjg <- Ounjg12 |> 
  rowSums()

# number of decorations as percentage
Ounjg12 <- Ounjg12 |> 
    mutate_if(is.numeric, function(x) x/n_ounjg*100)

names(Ounjg12)

n_ounjg <- data.frame(
  site = c("Ounjougou12"),
  n = c(n_ounjg)
)
```

## Sadia

The site was excavated as part of the "Human Population and Paleoenvironment in Africa" by Huysecom et al.. The dates of the different stratigraphic ensembles of the site are the results of a bayesian model based on 27 radiocarbon dates. For our 12^th^ century network we used pottery from the stratigraphic ensembles SE1 (1068-1209) and SE2 (1030-1133). In our 7^th^-9^th^ century network pottery from the earliest stratigraphic ensemble, SE5 (792-887), was included.

The ceramic data was taken from figure 9 of Huysecom et al. (2017).

It is problematic that for some SE the proportion of not specifally named decorations is high (especially for SE5 where this is the case for 25% of the decorations).

```{r}
Sadia <- read_csv2(here::here("analysis/data/raw_data/Sadia.csv")) |> 
  rename(straight_mat_impression = "mat impression",
         folded_strip_roulette = "cr4",
         braided_strip_roulette = "braided.strip",
         braided_cord_roulette = "braided.cord",
        unid_c_r_a = "unidentified composite roulette - A",
         unid_c_r_b = "unidentified composite roulette - B") |> 
  mutate(cord_wrapped_stick_roulette = unid_c_r_a + unid_c_r_b) |> 
  select(-plain, -unid_c_r_a, -unid_c_r_b)

#get total sherd number and percentage
n_sadia <- Sadia |> 
  select_if(is.numeric) |> 
  rowSums()

Sadia <- Sadia |> 
  add_column(n_sadia) |> 
  mutate_if(is.numeric, function(x) x/n_sadia*100)

#Selecting SE1 and SE2 for the 12th century network
Sadia12 <- Sadia |> 
  filter(Phase == "SE1" | Phase == "SE2") |> 
  summarise_if(is.numeric, mean) |> 
  select(-others,-n_sadia) |> 
  mutate(Site = "Sadia12") |> 
  column_to_rownames("Site")

#Selecting SE5 for the 9th century network
Sadia9 <- Sadia |> 
  filter(Phase == "SE5")|> 
  mutate(Site = "Sadia9") |> 
  column_to_rownames("Site")|> 
  select(-others, -Phase,-n_sadia)

#

names(Sadia)

n_sadia <- data.frame(
  site = c("Sadia12", "Sadia12", "Sadia9"),
  n = c(n_sadia[c(1, 2, 5)])
) |> 
  group_by(site) |> 
  summarise(n = sum(n))
```

# Collate datasets

Here we create two separate collated datasets for the 12th and 9th century networks by binding together the previously created data for each site and then further cleaning and standardising.

```{r 12c bind}
decors_12 <- bind_rows(Aku12_decor, AD12_decor, Dmssg12_decor, doupwil, essouk, galia, Gao12_decor, Jj12decor, Kkl12, mara_12, Ounjg12, Sadia12, shoma_12, tmd12_decor, togu_decor, tellem_decors) |> 
 mutate_all(~ifelse(is.na(.), 0, .)) |> 
  mutate(incised = incised + incised_patterns,
         cord_wrapped_roulette = cord_wrapped_roulette + cord_wrapped_stick_roulette,
         applied_plastic = applied_plastic + applied_ridges,
         punctate = punctate + impressed_stick,
         comb_impression = impressed_comb) |> 
  select(-c(incised_patterns,
            cord_wrapped_stick_roulette,
            slip, # suffers from uneven reporting
            unidentified_roulette,
            nonparallel_grooves,
            finger_arcs,
            applied_ridges,
            impressed_stick,
            impressed_comb)) |> 
  rownames_to_column("Site")

```

```{r 9c bind}
decors_9 <- bind_rows(Aku9_decor, Gao9_decor, Jj9decor, mara_9, Nin_Bere9, Sadia9, tmd9_decor, mdb_decor)|> 
  mutate_all(~ifelse(is.na(.), 0, .))|> 
  mutate(applied_plastic = applied_plastic + applied_ridges,
         cord_wrapped_roulette = cord_wrapped_roulette + cord_wrapped_stick_roulette) |> 
  select(-c(slip,
            unidentified_roulette,
            nonparallel_grooves,
            applied_ridges,
            cord_wrapped_stick_roulette)) |> 
  rownames_to_column("Site")

```

```{r}
n_decor <- rbind(n_akumbu, n_AD, n_Dmssg, n_doupwil, n_ess, n_galia, n_gao, n_jj, n_kkl, n_mara, n_mdb, n_NB, n_ounjg, n_sadia, n_sanga, n_Shoma12, n_tmd, n_togu)
```

# Saving data

We create two csv datasets to be used in further analyses.

```{r}
#12th century decor data
decors_12 |> 
write.csv2(here::here("analysis/data/derived_data/decors12.csv"))

#9th century decor data
decors_9 |> 
write.csv2(here::here("analysis/data/derived_data/decors9.csv"))

n_decor |> 
  write.csv2(here::here("analysis/data/derived_data/n_decor.csv"))

```